# Manuscript

```{r}
#| label: setup
#| include: false

source(here::here("R/quarto-setup.R"))
```

```{r}
#| echo: false
#| output: asis

rutils:::quarto_status(
  type = "drafting",
  of_what = "of this report"
  )
```

::: {.callout-note}
## Target journal

1. [Scientific Reports](https://www.nature.com/srep/author-instructions) ([IF 2022: 4.6/JCR](https://jcr.clarivate.com/jcr/) | [A1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
:::

::: {.callout-note}
The following study was performed by Arthur Ramalho Magalhães (__AM__) Daniel Vartanian (__DV__), and Aline Martins de Carvalho (__AC__).

__AC__ contributed to the design and implementation of the study. __DV__ and __MP__ collected the data and performed the statistical analysis. __AM__ wrote the manuscript. All authors discussed the results and revised the final manuscript.

_Future reference_: Magalhães, A. R., Vartanian, D., & Carvalho, A. M. (2024). A Some title. _Scientific Reports_.
:::
